---
title: "rstudio::conf 2017 recap"
author: "vp nagraj"
date: "2/16/2017"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
---
class: middle, center
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## <https://github.com/rstudio/rstudio-conf>

---

<h2> #1: tidyverse as "dialect"</h2>

```{r results='hide', warning = F, message = F}
{{ library(tidyverse) }} 
```

.pull-left[

### "regular R"

1. `subset()`

1. `read.csv()`

1. baz(bar(foo))

1. data frame]

.pull-right[

### "tidyverse R"

1. `filter()`

1. `read_csv()`

1. `foo %>% bar() %>% baz()`

1. tibble]

---
<h2> #2: list-cols</h2>
```{r, warning = FALSE, message = F}
library(gapminder)

gapminder %>%
  group_by(country) %>%
  nest() # from tidyr package
```

---

<h2> #3: purrr</h2>

<http://bit.ly/purrr-rstudioconf>

<https://jennybc.github.io/purrr-tutorial/>

**iterate a function over a list**

- `map()`: returns a list
- `map2()`: same as above but can iterate over two lists
- `map_chr()`: returns a character vector
- `map_lgl()`: returns a logical vector
- `map_dbl()`: returns a double precison (numeric) vector
- `map_int()`: returns an integer vector

---

<h2> #4: repurrrsive</h2>
```{r, message = FALSE, warning=FALSE}
library(purrr)
library(repurrrsive)

# how many colors are in each wes anderson movie palette?
wesanderson %>%
  map_int(length)

```
---

<h2> #4: repurrrsive</h2>
```{r, message = FALSE, warning=FALSE}

# which characters are included in game of thrones dataset?
got_chars %>%
  map_chr(~ .x$name)

```
---
<h2> #4: repurrrsive</h2>
```{r, message = FALSE, warning=FALSE}

got_chars %>%
  set_names(map_chr(got_chars, "name")) %>%
  map_lgl(~ .x$alive) %>%
  enframe()

```
---

<h2> #4: repurrrsive</h2>
```{r, warning = FALSE, message = F}

# note gap_simple is repurrrsive version of gapminder
gap_simple %>%
  group_by(country) %>%
  nest() %>%
  mutate(fit = map(data, ~ lm(lifeExp ~ year, data = .x)))

```
---
<h2> #5: fivevethirtyeight</h2>

<https://cran.r-project.org/web/packages/fivethirtyeight/vignettes/fivethirtyeight.html>

- daily_show_guests
- fandango
- love_actually_adj
- pulitzer
- tennis_players_time
- weather_check
- hist_ncaa_bball_casts
- congress_age
- airline_safety
- trump_twitter
- unisex_names
- tarantino
- nutrition_pvalues
- *etc*
---
<h2> #5: fivevethirtyeight</h2>

http://fivethirtyeight.com/features/a-statistical-analysis-of-the-work-of-bob-ross/

```{r, eval = F}
library(fivethirtyeight)
library(tidyverse)
library(ggthemes)

bob_ross %>%
  gather(feature, presence, apple_frame:wood_framed) %>%
  group_by(feature) %>%
  summarise(n = sum(presence), 
            perc = round(n / nrow(bob_ross), 2)) %>%
  filter(perc >= .02) %>%
  mutate(perc = paste0(perc*100, "%")) %>%
  arrange(desc(n)) %>%
  ggplot(aes(reorder(feature,n), n, label = perc)) +
  geom_bar(stat = "identity", fill = "dodgerblue3") +
  geom_text(nudge_y = 12) +
  coord_flip() +
  theme_fivethirtyeight()
```
---
<h2> #5: fivevethirtyeight</h2>
```{r, echo = F}
library(fivethirtyeight)
library(tidyverse)
library(ggthemes)

bob_ross %>%
  gather(feature, presence, apple_frame:wood_framed) %>%
  group_by(feature) %>%
  summarise(n = sum(presence), 
            perc = round(n / nrow(bob_ross), 2)) %>%
  filter(perc >= .02) %>%
  mutate(perc = paste0(perc*100, "%")) %>%
  arrange(desc(n)) %>%
  ggplot(aes(reorder(feature,n), n, label = perc)) +
  geom_bar(stat = "identity", fill = "dodgerblue3") +
  geom_text(nudge_y = 12) +
  coord_flip() +
  theme_fivethirtyeight()
```

---

<h2> #6: profvis</h2>

<https://rstudio.github.io/profvis/>

```{r, eval = FALSE}
mylm <- function(dat, pred, resp) {
  
  stopifnot(inherits(dat, "data.frame"))
  
  if (is.tibble(dat)) {
    
    dat <- as.data.frame(dat)
    
  }
  
  fit <- lm(dat[,pred] ~ dat[,resp])
  
  cat("coefficients are ...\n",fit$coefficients)
  
  plot(dat[,pred], dat[,resp])
  
}
```

---

<h2> #6: profvis</h2>

<https://rstudio.github.io/profvis/>

```{r, eval = FALSE}
# run these lines
# use keyboard shortcut (option + command + shift + p on MAC) to run and profile the lines
mylm(dat = mtcars, pred = "wt", resp = "mpg")
mylm(dat = congress_age, pred = "congress", resp = "age")

# alterantively wrap in profvis() function
profvis(mylm(dat = mtcars, pred = "wt", resp = "mpg"))
profvis(mylm(dat = congress_age, pred = "congress", resp = "age"))

```

---

<h2> #7: bookdown</h2>

<https://bookdown.org/yihui/bookdown/>

<https://ismayc.github.io/moderndiver-book/>

<http://nagraj.net/bims8382-textbook/>

<https://github.com/vpnagraj/bims8382-textbook>

---
<h2> #8: threejs</h2>

<http://bwlewis.github.io/rthreejs/>

```{r, echo = FALSE, warning=FALSE, message = FALSE}
library(outbreaks)
library(igraph)
library(threejs)

hospvisit <- 
  mers.korea.2015$contacts %>%
  filter(exposure == "Visit hospital") %>%
  select(from,to)

hospvisitll <-
  mers.korea.2015$linelist %>%
  filter(id %in% hospvisit$from | id %in% hospvisit$to) %>%
  select(id)
  
# create graph object from data frame
g <- graph_from_data_frame(hospvisit, vertices = hospvisitll)

# remove edges from graph
gnoe <- g - edges(1:12)

# make list of graphs
graph_list <- list(gnoe,
                   gnoe + edge(ends(g, es = E(g), names = FALSE)[1,]), 
                   gnoe + edge(ends(g, es = E(g), names = FALSE)[1,]) + edge(ends(g, es = E(g), names = FALSE)[2,]),
                   gnoe + edge(ends(g, es = E(g), names = FALSE)[1,]) + edge(ends(g, es = E(g), names = FALSE)[2,]) +  edge(ends(g, es = E(g), names = FALSE)[3,]),
                   gnoe + edge(ends(g, es = E(g), names = FALSE)[1,]) + edge(ends(g, es = E(g), names = FALSE)[2,]) +  edge(ends(g, es = E(g), names = FALSE)[3,]) + edge(ends(g, es = E(g), names = FALSE)[4,])
                   )

# plot graphs                   
graphjs(graph_list,
        vertex.shape="sphere", 
        vertex.label = names(V(gnoe)),
        edge.width=3)
```

---

<h2> #8: threejs</h2>

```{r, eval = FALSE}
library(threejs)
library(outbreaks)

# prep edge attributes
hospvisit <- 
  mers.korea.2015$contacts %>%
  filter(exposure == "Visit hospital") %>%
  select(from,to)

# prep node attributes
hospvisitll <-
  mers.korea.2015$linelist %>%
  filter(id %in% hospvisit$from | id %in% hospvisit$to) %>%
  select(id)
  
# create graph object from data frame
g <- graph_from_data_frame(hospvisit, vertices = hospvisitll)

# remove edges from graph
gnoe <- g - edges(1:12)
```

---

<h2> #8: threejs</h2>

<https://bwlewis.github.io/rthreejs/animation/animation.html>

```{r, eval = FALSE}
# make list of graphs
graph_list <- list(gnoe,
                   gnoe + edge(ends(g, es = E(g), names = FALSE)[1,]), 
                   gnoe + edge(ends(g, es = E(g), names = FALSE)[1,]) + edge(ends(g, es = E(g), names = FALSE)[2,]),
                   gnoe + edge(ends(g, es = E(g), names = FALSE)[1,]) + edge(ends(g, es = E(g), names = FALSE)[2,]) +  edge(ends(g, es = E(g), names = FALSE)[3,]),
                   gnoe + edge(ends(g, es = E(g), names = FALSE)[1,]) + edge(ends(g, es = E(g), names = FALSE)[2,]) +  edge(ends(g, es = E(g), names = FALSE)[3,]) + edge(ends(g, es = E(g), names = FALSE)[4,])
                   )

# plot graphs                   
graphjs(graph_list,
        vertex.shape="sphere", 
        vertex.label = names(V(gnoe)),
        edge.width=3)

```
---
<h2> #8: threejs</h2>
```{r, eval = FALSE}
library(threejs)
g <- make_ring(5) - edges(1:5) # five vertices, no edges

graph_list <- list(
  g + edge(1, 2),
  g + edge(1, 2) + edge(2, 3),
  g + edge(1, 2) + edge(2, 3) + edge(3, 4),
  g + edge(1, 2) + edge(2, 3) + edge(3, 4) + edge(4, 5),
  g + edge(1, 2) + edge(2, 3) + edge(3, 4) + edge(4, 5) + edge(5, 1))

graphjs(graph_list, main=paste(1:5),
        vertex.color=rainbow(5), 
        vertex.shape="sphere", 
        edge.width=3)
```

---

<h2> #8: threejs</h2>

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(threejs)
g <- make_ring(5) - edges(1:5) # five vertices, no edges

graph_list <- list(
  g + edge(1, 2),
  g + edge(1, 2) + edge(2, 3),
  g + edge(1, 2) + edge(2, 3) + edge(3, 4),
  g + edge(1, 2) + edge(2, 3) + edge(3, 4) + edge(4, 5),
  g + edge(1, 2) + edge(2, 3) + edge(3, 4) + edge(4, 5) + edge(5, 1))

graphjs(graph_list, main=paste(1:5),
        vertex.color=rainbow(5), vertex.shape="sphere", edge.width=3)
```

---

<h2> #9: xaringan</h2>

<http://slides.yihui.name/xaringan/>

## Highlighting Code

.pull-left[

    `r ''````
    nba_tattoos %>%
      group_by(tattoos) %>%
      summarise(n = n()) %>%
      ggplot(aes(tattoos, n)) +
      geom_bar(stat = "identity") +
    {{ theme_excel() }}
    ```
]
.pull-right[

```{r tidy=FALSE, results = "hide", eval = FALSE}
nba_tattoos %>%
  group_by(tattoos) %>%
  summarise(n = n()) %>%
  ggplot(aes(tattoos, n)) +
  geom_bar(stat = "identity") +
{{ theme_excel() }}
```

]

---

<h2> #9: xaringan</h2>

```{r echo = FALSE}
nba_tattoos %>%
  group_by(tattoos) %>%
  summarise(n = n()) %>%
  ggplot(aes(tattoos, n)) +
  geom_bar(stat = "identity") +
  theme_excel()
```

---
<h2> #10: everything else</h2>

- **sparklyr**: R interface for using **dplyr** syntax in Spark (big data) queries

  <http://spark.rstudio.com/>
  
- **shinytest**: record expected behavior of a **shiny** app to benchmark performance 

  <https://rstudio.github.io/shinytest/>
  
- **flexdashboard**: dashboard output format for R Markdown documents

  <http://rmarkdown.rstudio.com/flexdashboard/index.html>

- **rcpp**: integrate C++ code with R

  <http://www.rcpp.org/>

---
class: middle

## <https://www.rstudio.com/resources/webinars/#rstudioconf>
